---
title: "Diddy Visualizations"
output: html_document
date: "2024-11-01"
---

# Initial Visualizations

```{r}
library(tidyverse)
library(forcats)
library(patchwork)
```

```{r}
diddy_csv <- read_csv("~/BSAFall2024/DiddyProject/Diddy-Stats/diddygames.csv")
```

```{r}
diddy_csv <- diddy_csv |> 
  select(-`...1`)
```

```{r}
diddy_csv |> head()
```

```{r}
diddy_csv <- diddy_csv |> filter(mp != "Did Not Dress")
```

### Team Colors

<https://nbacolors.com/team/miami-heat-colors>

```{r}
team_colors <- c(
  "MIA" = "#98002e",
  "BOS" = "#007a33",
  "NJN" = "#c6cfd4",
  "BRK" = "black",
  "CHI" = "#ce1141",
  "PHO" = "#e56020",
  "DEN" = "#0e2240",
  "GSW" = "#1d428a",
  "IND" = "#002d62",
  "MIL" = "#00471b",
  "NOP" = "#85714d",
  "NOH" = "#17bfdd",
  "OKC" = "#007ac1",
  "PHI" = "#006bb6",
  "SAC" = "#5a2d81",
  "TOR" = "#ba0c2f",
  "CLE" = "#860038",
  "LAL" = "#552583",
  "LAC" = "#c8102e",
  "CHO" = "#00788c",
  "ATL" = "#e03a3e",
  "DAL" = "#00538c",
  "DET" = "#1d42ba",
  "HOU" = "#ffc72c",
  "MEM" = "#5d76a9",
  "MIN" = "#78be20",
  "NYK" = "#f58426",
  "ORL" = "#0077c0",
  "POR" = "#e03a3e",
  "SAS" = "#8a8d8f",
  "UTA" = "#f9a01b",
  "WAS" = "#002b5c"
)
```

## Getting distribution of teams

Diddy went to 31 Games, which were the teams he watched the most often

```{r}
game_count_df <- diddy_csv |>
  group_by(date, team) |>
  slice(1) |>
  ungroup()
```

```{r}
game_count_df
```

```{r}
ggplot(game_count_df, aes(x = fct_infreq(team), fill = team)) + 
  geom_bar() + 
  scale_fill_manual(values = team_colors) + 
  labs(title="Diddy's Favorite Teams to Watch", x="Teams", y="Number of Games with Diddy in Attendance") +
  guides(fill="none") +
  theme_classic()
```

## Getting the Distribution of Players

```{r}
unique(diddy_csv$player) |> length()
```

```{r}
player_count_df <- diddy_csv |> 
  count(player, sort = TRUE) |>
  slice_head(n = 20)

player_count_df$player
```

```{r}
top_20_players <- diddy_csv |>
  filter(player %in% player_count_df$player) |> 
  arrange(player)
```

```{r}
ggplot(top_20_players, aes(x = fct_infreq(player), fill = team)) +
  geom_bar() + 
  scale_fill_manual(values = team_colors) + 
  labs(title="The Top 20 Players with the Most Diddy Games",x="Player", y="Number of Games with Diddy in Attendance") +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Stats comparison for each player

## Adding columns for comparison

### Playoffs vs Regular season

```{r}
playoff_dates <- as.Date(c("2009-04-27", "2023-04-24", "2013-04-23"))
```

```{r}
top_20_playoffs <- top_20_players |> 
  mutate(game_type = if_else(
    month(date) %in% c(5, 6)| date %in% playoff_dates,
    "playoff",
    "regular"
  ))
```

```{r}
top_20_playoffs
```

### Adding season

```{r}
top_20_playoffs_seasons <- top_20_playoffs |> 
  mutate(season = case_when(
    date >= as.Date("1996-10-01") & date <= as.Date("1997-07-31") ~ "1996-97",
    date >= as.Date("1997-10-01") & date <= as.Date("1998-07-31") ~ "1997-98",
    date >= as.Date("1998-10-01") & date <= as.Date("1999-07-31") ~ "1998-99",
    date >= as.Date("1999-10-01") & date <= as.Date("2000-10-11") ~ "1999-00",
    date >= as.Date("2000-10-01") & date <= as.Date("2001-07-31") ~ "2000-01",
    date >= as.Date("2001-12-01") & date <= as.Date("2002-07-31") ~ "2001-02",
    date >= as.Date("2002-10-01") & date <= as.Date("2003-07-31") ~ "2002-03",
    date >= as.Date("2003-10-01") & date <= as.Date("2004-07-31") ~ "2003-04",
    date >= as.Date("2004-10-01") & date <= as.Date("2005-07-31") ~ "2004-05",
    date >= as.Date("2005-10-01") & date <= as.Date("2006-07-31") ~ "2005-06",
    date >= as.Date("2006-10-01") & date <= as.Date("2007-07-31") ~ "2006-07",
    date >= as.Date("2007-10-01") & date <= as.Date("2008-07-31") ~ "2007-08",
    date >= as.Date("2008-10-01") & date <= as.Date("2009-07-31") ~ "2008-09",
    date >= as.Date("2009-10-01") & date <= as.Date("2010-10-11") ~ "2009-10",
    date >= as.Date("2010-10-01") & date <= as.Date("2011-07-31") ~ "2010-11",
    date >= as.Date("2011-12-01") & date <= as.Date("2012-07-31") ~ "2011-12",
    date >= as.Date("2012-10-01") & date <= as.Date("2013-07-31") ~ "2012-13",
    date >= as.Date("2013-10-01") & date <= as.Date("2014-07-31") ~ "2013-14",
    date >= as.Date("2014-10-01") & date <= as.Date("2015-07-31") ~ "2014-15",
    date >= as.Date("2015-10-01") & date <= as.Date("2016-07-31") ~ "2015-16",
    date >= as.Date("2016-10-01") & date <= as.Date("2017-07-31") ~ "2016-17",
    date >= as.Date("2017-10-01") & date <= as.Date("2018-07-31") ~ "2017-18",
    date >= as.Date("2018-10-01") & date <= as.Date("2019-07-31") ~ "2018-19",
    date >= as.Date("2019-10-01") & date <= as.Date("2020-10-11") ~ "2019-20",
    date >= as.Date("2020-12-01") & date <= as.Date("2021-07-31") ~ "2020-21",
    date >= as.Date("2021-10-01") & date <= as.Date("2022-07-31") ~ "2021-22",
    date >= as.Date("2022-10-01") & date <= as.Date("2023-07-31") ~ "2022-23",
    date >= as.Date("2023-10-01") & date <= as.Date("2024-07-31") ~ "2023-24",
    TRUE ~ "Other"
  ))
```

```{r}
top_20_playoffs_seasons
```

### Importing playoff and regular season stats

```{r}
career_stats_df <- read_csv("~/BSAFall2024/DiddyProject/Diddy-Stats/top20.csv")
```

```{r}
colnames(career_stats_df)
```

```{r}
career_stats_df <- career_stats_df |> 
  select(-c(pos, lg_id, team_id, gs, `...1`, age, DUMMY, trp_dbl))
```

```{r}
career_stats_df 
```

```{r}
career_stats_pg_df <- career_stats_df |>
  mutate(across(c(mp, fg, fga, fg3, fg3a, fg2, fg2a, ft, fta, orb, drb, trb, ast, stl,
                  blk, tov, pf, pts), function(x) round(x / g, digits = 2)))
```

```{r}
career_stats_pg_df
```

For each player I want to see how their games compare to their averages for that season in the playoffs/regular season

```{r}
combine_career_diddy <- function(diddy_df, career_df, player_name) {
  diddy <- diddy_df |> 
    filter(tolower(player) == tolower(player_name)) |>
    rename(season_type = game_type)
  
  career <- career_df |>
    filter(tolower(player) == tolower(player_name)) |>
    select(-player)
  
  combined_df <- diddy |>
    left_join(
      career |> rename_with(~paste0("career_", .x), -c(season, season_type)),
      by = c("season", "season_type")
    )
  
  return(combined_df)
}
```

```{r}
combined_plots <- function(combined_df, stat) { 
  player_name <- combined_df$player[1]
  select_cols <- c("player", "season", "season_type", "date", stat, paste0("career_", stat))
  
  combined_df <- combined_df |>
    select(all_of(select_cols))
  
  combined_df_long <- combined_df |> 
    pivot_longer(cols = all_of(c(stat, paste0("career_", stat))),
               names_to = "stat_type",
               values_to = "value") |>
    mutate(stat_type = case_when(
      stat_type == paste0("career_", stat) ~ paste("Season", stat),
      stat_type == stat ~ paste("Diddy", stat)
    ))
  
  combined_df_long <- combined_df |> 
    pivot_longer(cols = all_of(c(stat, paste0("career_", stat))),
                 names_to = "stat_type",
                 values_to = "value") |>
    mutate(stat_type = case_when(
      stat_type == paste0("career_", stat) ~ paste("Season", stat),
      stat_type == stat ~ paste("Diddy", stat)
    ))
  
  fill_labels <- setNames(c("red", "green"), 
                          c(paste0("Diddy ", stat), paste0("Season ", stat)))
  
  ggplot(combined_df_long, 
         aes(x = factor(date), y = value, fill = stat_type)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    scale_fill_manual(values = fill_labels) +
    facet_wrap(~season_type, nrow = 2) +
    labs(x = "Date of Game", y = stat, title=paste0("Diddy's effect on ", player_name, "'s ", stat), fill = "Stat Type") + 
    theme_bw()
}
```

```{r}
stat_differential <- function(combined_df) {
  # c(pts, pf, tov, blk, stl, ast, trb)
  
  diff_df <- combined_df |> 
    mutate(diff_pts = pts - career_pts,
           diff_pf = pf - career_pf,
           diff_tov = tov - career_tov,
           diff_blk = blk - career_blk,
           diff_stl = stl - career_stl, 
           diff_ast = ast - career_ast,
           diff_trb = trb - career_trb,
             ) |> 
    select(
          player, season, season_type, diff_pts, diff_ast, diff_trb, diff_tov, diff_blk, diff_stl, diff_pf
      )
  
  return(diff_df)
}

```

```{r}
# take output of stat differential and get the averages grouped by season_type
diff_averages <- function(stat_diff_df) {
  player <- stat_diff_df$player[1]
  
  diff_avgs <- stat_diff_df |> group_by(season_type) |>
      summarize(across(starts_with('diff_'), mean))
  
  diff_avgs$player <- player
  
  diff_avgs <- diff_avgs |> select(player, everything())
  
  return(diff_avgs)
}
```

## Dwyane Wade

```{r}
top_20_playoffs_seasons
```

```{r}
dwade <- "Dwyane Wade"

dwade_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, dwade) 
```

```{r}
dwade_combined_df
```

```{r}
dwade_pts <- combined_plots(dwade_combined_df, "pts")
dwade_ast <- combined_plots(dwade_combined_df, "ast")
dwade_trb <- combined_plots(dwade_combined_df, "trb")
```

```{r}
dwade_pts
```

```{r}
dwade_ast
```

```{r}
dwade_trb
```

```{r}
dwade_stat_diff <- stat_differential(dwade_combined_df)

dwade_stat_diff
```

```{r}
diff_averages(dwade_stat_diff)
```

## Kobe Bryant

```{r}
kobe <- "Kobe Bryant"

kobe_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, kobe)
```

```{r}
kobe_combined_df
```

```{r}
kobe_pts <- combined_plots(kobe_combined_df, "pts")
kobe_ast <- combined_plots(kobe_combined_df, "ast")
kobe_trb <- combined_plots(kobe_combined_df, "trb")
```

```{r}
kobe_pts
```

```{r}
kobe_ast
```

```{r}
kobe_trb
```

```{r}
kobe_stat_diff <- stat_differential(kobe_combined_df)
```

```{r}
kobe_stat_diff
```

## Lamar Odom

```{r}
lodom <- "Lamar Odom"

lodom_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, lodom)
lodom_pts <- combined_plots(lodom_combined_df, "pts")
lodom_ast <- combined_plots(lodom_combined_df, "ast")
lodom_trb <- combined_plots(lodom_combined_df, "trb")
```

```{r}
lodom_pts
lodom_ast
lodom_trb
```

```{r}
stat_differential(lodom_combined_df)
```

## Pau Gasol

```{r}
gasol <- "Pau Gasol"

gasol_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, gasol)
gasol_pts <- combined_plots(gasol_combined_df, "pts")
gasol_ast <- combined_plots(gasol_combined_df, "ast")
gasol_trb <- combined_plots(gasol_combined_df, "trb")

gasol_diff <- stat_differential(gasol_combined_df)
```


```{r}
colnames(career_stats_df)

gasol_career_totals <- career_stats_df |> filter(player == "Pau Gasol") |> 
  select(pts, ast, trb, tov,
                       blk, stl, pf, g) |>
  colSums()
  
gasol_career_totals
```
```{r}
gasol_career_totals / gasol_career_totals[["g"]]
```


```{r}
gasol_combined_df |> 
  select(pts, ast, trb, tov,
                       blk, stl, pf) |>
  colSums() / 8 
```

```{r}
gasol_diff
```

```{r}
gasol_diff |> 
  group_by(season_type) |>
  summarize(across(c(diff_pts, diff_ast, diff_trb, diff_tov,
                       diff_blk, diff_stl, diff_pf), mean), .groups = "drop")

gasol_diff |> 
  summarize(across(c(diff_pts, diff_ast, diff_trb, diff_tov,
                       diff_blk, diff_stl, diff_pf), mean), .groups = "drop")
```

## Derek Fisher

## Lebron James

```{r}
goat <- "LeBron James"

goat_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, goat)
goat_pts <- combined_plots(goat_combined_df, "pts")
goat_ast <- combined_plots(goat_combined_df, "ast")
goat_trb <- combined_plots(goat_combined_df, "trb")

goat_diff <- stat_differential(goat_combined_df)
```

```{r}
goat_diff |> 
  group_by(season_type) |>
  summarize(across(c(diff_pts, diff_ast, diff_trb, diff_tov,
                       diff_blk, diff_stl, diff_pf), mean), .groups = "drop")

goat_diff |> 
  summarize(across(c(diff_pts, diff_ast, diff_trb, diff_tov,
                       diff_blk, diff_stl, diff_pf), mean), .groups = "drop")
```

```{r}
goat_pts
goat_ast
goat_trb
```

## Jordan Farmar

## Sasha Vujačić

## Trevor Ariza

## Udonis Haslem

## Andrew Bynum

## Jason Kidd

## Kevin Garnett

## Luke Walton

```{r}
walton <- "Luke Walton"

walton_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, walton)
```

```{r}
walton_combined_df
```

```{r}

walton_diddy_minutes <- sapply(walton_combined_df$mp , function(time) {
  parts <- as.numeric(unlist(strsplit(time, ":")))
  hours <- parts[1]
  minutes <- parts[2]
  hours + minutes / 60
}) |> unname()



mean(walton_combined_df$career_mp)

mean((walton_combined_df$career_mp - walton_diddy_minutes) / (walton_combined_df$career_mp))
```

## Paul Pierce

```{r}
ppierce <- "Paul Pierce"

ppierce_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, ppierce)
ppierce_pts <- combined_plots(ppierce_combined_df, "pts")
ppierce_ast <- combined_plots(ppierce_combined_df, "ast")
ppierce_trb <- combined_plots(ppierce_combined_df, "trb")

stat_differential(ppierce_combined_df)
```

```{r}

```

## Rajon Rondo

## Ray Allen

## Shannon Brown

## J.R. Smith

## Chris Bosh

## Everybody combined

```{r}
get_diff_all <- function(diddy_df, career_df) {
  players <- diddy_df$player |> unique() |> as.list()
  
  combined_df_list <- lapply(players, function(name) {
    combine_career_diddy(diddy_df, career_df, name)
  })
  
  stat_diff_list <- lapply(combined_df_list, function(df) stat_differential(df))
  
  avg_stat_diff_list <- lapply(stat_diff_list, function(df) diff_averages(df))
  
  return(bind_rows(avg_stat_diff_list))
}
```

```{r}
diff_all_df <- get_diff_all(top_20_playoffs_seasons, career_stats_pg_df)
```

```{r}
ggplot(diff_all_df, aes(x = player, y = diff_pts, fill = diff_pts > 0)) +
  geom_col() +
  scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
  labs(x = "Players", y = "Diddy Point Differential") +
  facet_wrap(~season_type, nrow = 2) + 
  theme_bw() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
  
```

```{r}
ggplot(diff_all_df, aes(x = player, y = diff_ast, fill = diff_ast > 0)) +
  geom_col() +
  scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
  labs(x = "Players", y = "Diddy Assist Differential") +
  facet_wrap(~season_type, nrow = 2) + 
  theme_bw() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
  
```

```{r}
ggplot(diff_all_df, aes(x = player, y = diff_trb, fill = diff_trb > 0)) +
  geom_col() +
  scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
  labs(x = "Players", y = "Diddy Rebound Differential") +
  facet_wrap(~season_type, nrow = 2) + 
  theme_bw() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
  
```

```{r}
all_positive_df <- diff_all_df |> 
  filter(diff_ast > 0 & diff_pts > 0 & diff_trb > 0)
```

```{r}
all_positive_df
```

```{r}
summary(c(7, 4, 6, 3, 4, 4, 5))
```

## Tangent into associations between PRA and TOs

```{r}
pp_playoff_df <- read_csv("~/BSAFall2024/DiddyProject/career_data/paul_pierce_playoffs.csv")

pp_playoff_df <- pp_playoff_df |> select(-`...1`)
```

Creating the PRA stat

```{r}
pp_playoff_df <- pp_playoff_df |> 
  mutate(pra = pts + trb + ast)
```

```{r}
pp_playoff_df
```

```{r}
pp_lm <- lm(tov~pra, data = pp_playoff_df)
```

```{r}
summary(pp_lm)
```

```{r}
ppierce_combined_df$date
```

```{r}
pp_playoff_df <- pp_playoff_df |> 
  mutate(diddy = if_else(date_game %in% ppierce_combined_df$date, T, F))
```

```{r}
ggplot(pp_playoff_df, aes(x = pra, y = tov)) + 
  geom_point(aes(color = diddy)) + 
  geom_smooth(method="lm", color = "black", alpha = 0.7) + 
  labs(y = "Turnovers", x = "Points, Rebounds, and Assists (PRA)", title = "Relationship between PRA and TOV for Paul Pierce in the Playoffs", color = "Diddy Attendance") +
  theme_classic()
```

```{r}
wade_regular_df <- read_csv("~/BSAFall2024/DiddyProject/career_data/dwyane_wade_regular.csv")
```

```{r}
wade_regular_df <- wade_regular_df |>
  mutate(pra = pts + trb + ast)
```

```{r}
wade_regular_df <- wade_regular_df |> 
  mutate(diddy = if_else(date_game %in% dwade_combined_df$date, T, F))
```

```{r}
ggplot(wade_regular_df, aes(x = pra, y = tov)) + 
  geom_point(aes(color = diddy)) + 
  geom_smooth(method="lm", color = "black", alpha = 0.7) + 
  labs(y = "Turnovers", x = "Points, Rebounds, and Assists (PRA)", title = "Relationship between PRA and TOV for Dwyane Wade in the Regular Season", color = "Diddy Attendance") +
  theme_classic()
```

```{r}
dwade_lm <- lm(tov~pra, data = wade_regular_df)
```

```{r}
summary(dwade_lm)
```

## All Negative

```{r}
all_negative_df <- diff_all_df |> 
  filter(diff_ast < 0 & diff_pts < 0 & diff_trb < 0)
```

```{r}
all_negative_df
```
