---
title: "Diddy Visualizations"
output: html_document
date: "2024-11-01"
---

# Initial Visualizations

```{r}
library(tidyverse)
library(forcats)
library(patchwork)
```

```{r}
diddy_csv <- read_csv("~/BSAFall2024/DiddyProject/Diddy-Stats/diddygames.csv")
```

```{r}
diddy_csv <- diddy_csv |> 
  select(-`...1`)
```

```{r}
diddy_csv |> head()
```

```{r}
diddy_csv <- diddy_csv |> filter(mp != "Did Not Dress")
```

```{r}
diddy_csv <- diddy_csv |> 
  mutate(player = if_else(player == "Dwyane Wade", "Dwayne Wade", player))
```

### Team Colors

<https://nbacolors.com/team/miami-heat-colors>

```{r}
team_colors <- c(
  "MIA" = "#98002e",
  "BOS" = "#007a33",
  "NJN" = "#c6cfd4",
  "BRK" = "black",
  "CHI" = "#ce1141",
  "PHO" = "#e56020",
  "DEN" = "#0e2240",
  "GSW" = "#1d428a",
  "IND" = "#002d62",
  "MIL" = "#00471b",
  "NOP" = "#85714d",
  "NOH" = "#17bfdd",
  "OKC" = "#007ac1",
  "PHI" = "#006bb6",
  "SAC" = "#5a2d81",
  "TOR" = "#ba0c2f",
  "CLE" = "#860038",
  "LAL" = "#552583",
  "LAC" = "#c8102e",
  "CHO" = "#00788c",
  "ATL" = "#e03a3e",
  "DAL" = "#00538c",
  "DET" = "#1d42ba",
  "HOU" = "#ffc72c",
  "MEM" = "#5d76a9",
  "MIN" = "#78be20",
  "NYK" = "#f58426",
  "ORL" = "#0077c0",
  "POR" = "#e03a3e",
  "SAS" = "#8a8d8f",
  "UTA" = "#f9a01b",
  "WAS" = "#002b5c"
)
```

## Getting distribution of teams

Diddy went to 31 Games, which were the teams he watched the most often

```{r}
game_count_df <- diddy_csv |>
  group_by(date, team) |>
  slice(1) |>
  ungroup()
```

```{r}
game_count_df
```

```{r}
ggplot(game_count_df, aes(x = fct_infreq(team), fill = team)) + 
  geom_bar() + 
  scale_fill_manual(values = team_colors) + 
  labs(title="Diddy's Favorite Teams to Watch", x="Teams", y="Number of Games with Diddy in Attendance") +
  guides(fill="none") +
  theme_classic()
```

## Getting the Distribution of Players

```{r}
player_count_df <- diddy_csv |> 
  count(player, sort = TRUE) |>
  slice_head(n = 20)

player_count_df$player
```

```{r}
top_20_players <- diddy_csv |>
  filter(player %in% player_count_df$player) |> 
  arrange(player)
```

```{r}
ggplot(top_20_players, aes(x = fct_infreq(player), fill = team)) +
  geom_bar() + 
  scale_fill_manual(values = team_colors) + 
  labs(title="The Top 20 Players with the Most Diddy Games",x="Player", y="Number of Games with Diddy in Attendance") +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Stats comparison for each player

## Adding columns for comparison

### Playoffs vs Regular season

```{r}
playoff_dates <- as.Date(c("2009-04-27", "2023-04-24", "2013-04-23"))
```

```{r}
top_20_playoffs <- top_20_players |> 
  mutate(game_type = if_else(
    month(date) %in% c(5, 6)| date %in% playoff_dates,
    "playoff",
    "regular"
  ))
```

```{r}
top_20_playoffs
```

### Adding season

```{r}
top_20_playoffs_seasons <- top_20_playoffs |> 
  mutate(season = case_when(
    date >= as.Date("1996-10-01") & date <= as.Date("1997-07-31") ~ "1996-97",
    date >= as.Date("1997-10-01") & date <= as.Date("1998-07-31") ~ "1997-98",
    date >= as.Date("1998-10-01") & date <= as.Date("1999-07-31") ~ "1998-99",
    date >= as.Date("1999-10-01") & date <= as.Date("2000-10-11") ~ "1999-00",
    date >= as.Date("2000-10-01") & date <= as.Date("2001-07-31") ~ "2000-01",
    date >= as.Date("2001-12-01") & date <= as.Date("2002-07-31") ~ "2001-02",
    date >= as.Date("2002-10-01") & date <= as.Date("2003-07-31") ~ "2002-03",
    date >= as.Date("2003-10-01") & date <= as.Date("2004-07-31") ~ "2003-04",
    date >= as.Date("2004-10-01") & date <= as.Date("2005-07-31") ~ "2004-05",
    date >= as.Date("2005-10-01") & date <= as.Date("2006-07-31") ~ "2005-06",
    date >= as.Date("2006-10-01") & date <= as.Date("2007-07-31") ~ "2006-07",
    date >= as.Date("2007-10-01") & date <= as.Date("2008-07-31") ~ "2007-08",
    date >= as.Date("2008-10-01") & date <= as.Date("2009-07-31") ~ "2008-09",
    date >= as.Date("2009-10-01") & date <= as.Date("2010-10-11") ~ "2009-10",
    date >= as.Date("2010-10-01") & date <= as.Date("2011-07-31") ~ "2010-11",
    date >= as.Date("2011-12-01") & date <= as.Date("2012-07-31") ~ "2011-12",
    date >= as.Date("2012-10-01") & date <= as.Date("2013-07-31") ~ "2012-13",
    date >= as.Date("2013-10-01") & date <= as.Date("2014-07-31") ~ "2013-14",
    date >= as.Date("2014-10-01") & date <= as.Date("2015-07-31") ~ "2014-15",
    date >= as.Date("2015-10-01") & date <= as.Date("2016-07-31") ~ "2015-16",
    date >= as.Date("2016-10-01") & date <= as.Date("2017-07-31") ~ "2016-17",
    date >= as.Date("2017-10-01") & date <= as.Date("2018-07-31") ~ "2017-18",
    date >= as.Date("2018-10-01") & date <= as.Date("2019-07-31") ~ "2018-19",
    date >= as.Date("2019-10-01") & date <= as.Date("2020-10-11") ~ "2019-20",
    date >= as.Date("2020-12-01") & date <= as.Date("2021-07-31") ~ "2020-21",
    date >= as.Date("2021-10-01") & date <= as.Date("2022-07-31") ~ "2021-22",
    date >= as.Date("2022-10-01") & date <= as.Date("2023-07-31") ~ "2022-23",
    date >= as.Date("2023-10-01") & date <= as.Date("2024-07-31") ~ "2023-24",
    TRUE ~ "Other"
  ))
```

```{r}
top_20_playoffs_seasons
```

### Importing playoff and regular season stats

```{r}
career_stats_df <- read_csv("~/BSAFall2024/DiddyProject/Diddy-Stats/top20.csv")
```

```{r}
colnames(career_stats_df)
```

```{r}
career_stats_df <- career_stats_df |> 
  select(-c(pos, lg_id, team_id, gs, `...1`, age, DUMMY, trp_dbl))
```

```{r}
career_stats_df 
```

```{r}
career_stats_pg_df <- career_stats_df |>
  mutate(across(c(mp, fg, fga, fg3, fg3a, fg2, fg2a, ft, fta, orb, drb, trb, ast, stl,
                  blk, tov, pf, pts), function(x) round(x / g, digits = 2)))
```

```{r}
career_stats_pg_df
```

For each player I want to see how their games compare to their averages for that season in the playoffs/regular season

```{r}
combine_career_diddy <- function(diddy_df, career_df, player_name) {
  diddy <- diddy_df |> 
    filter(player == player_name) |>
    rename(season_type = game_type)
  
  career <- career_df |>
    filter(player == player_name) |>
    select(-player)
  
  combined_df <- diddy |>
    left_join(
      career |> rename_with(~paste0("career_", .x), -c(season, season_type)),
      by = c("season", "season_type")
    )
  
  return(combined_df)
}
```

```{r}
combined_plots <- function(combined_df, stat) { 
  player_name <- combined_df$player[1]
  select_cols <- c("player", "season", "season_type", "date", stat, paste0("career_", stat))
  
  combined_df <- combined_df |>
    select(all_of(select_cols))
  
  combined_df_long <- combined_df |> 
    pivot_longer(cols = all_of(c(stat, paste0("career_", stat))),
               names_to = "stat_type",
               values_to = "value") |>
    mutate(stat_type = case_when(
      stat_type == paste0("career_", stat) ~ paste("Season", stat),
      stat_type == stat ~ paste("Diddy", stat)
    ))
  
  ggplot(combined_df_long, 
         aes(x = factor(date), y = value, fill = stat_type)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~season_type, nrow = 2) +
    labs(x = "Date of Game", y = stat, title=paste0("Diddy's effect on ", player_name, "'s ", stat)) + 
    theme_bw() + 
    theme(axis.text.x = element_blank())
}
```

## Dwayne Wade

```{r}
dwade <- "Dwayne Wade"

dwade_combined_df <- combine_career_diddy(top_20_playoffs_seasons, career_stats_pg_df, dwade)
```

```{r}
dwade_pts <- combined_plots(dwade_combined_df, "pts")
dwade_ast <- combined_plots(dwade_combined_df, "ast")
dwade_trb <- combined_plots(dwade_combined_df, "trb")
```

```{r}
dwade_pts
```

## Kobe Bryant

## Lamar Odom

## Pau Gasol

## Derek Fisher

## Lebron James

## Jordan Farmar

## Sasha Vujačić

## Trevor Ariza

## Udonis Haslem

## Andrew Bynum

## Jason Kidd

## Kevin Garnett

## Luke Walton

## Paul Pierce

## Rajon Rondo 

## Ray Allen

## Shannon Brown

## J.R. Smith

## Chris Bosh
